Short-term Path Planning with Collision Avoidance in Autonomous Driving based on Adaptive MPC

### 2019-02-04 Thesis_Latex/Paper/PaperFinalVersion.tex
update full version Paper

\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage{listings}
\usepackage{stfloats}
\usepackage{midfloat}
\usepackage{color}
\usepackage[table]{xcolor}
\usepackage{lipsum}
\usepackage[utf8x]{inputenc}

\usepackage{bbm}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[colorlinks=true,linkcolor=black,anchorcolor=black,citecolor=black,filecolor=black,menucolor=black,runcolor=black,urlcolor=black]{hyperref}

\newtheorem{thm}{Theorem}[section]
\newtheorem{defn}[thm]{Def.}
\newtheorem{propen}[thm]{Prop.}
\newtheorem{cor}[thm]{Cor.}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
		T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\title{Advanced Methods for short-term Path Planning in Autonomous Driving based on Adaptive MPC}
\author{\IEEEauthorblockN{Alberto Franco}\\
\IEEEauthorblockA{Department of Information Engineering\\
Università degli Studi di Padova\\ \href{mailto:alberto.franco.3@studenti.unipd.it}{alberto.franco.3@studenti.unipd.it}}}

\begin{document}
\maketitle
\begin{abstract}
	In this paper we present two different strategies for a self-driving car short-term path planning. The aim of this project is to study and implement a path tracking framework in order to prevent ATLASCAR2 collision with moving obstacle vehicles. The proposed algorithms, based on the Model Predictive Control paradigm, solve optimal problems formulated in terms of cost minimization under constraints.
\end{abstract}
\section{Introduction}
In robotic research, the problem of navigation is among the most important. Basically all autonomous mobile robots need some kind of navigation to fulfill the mobile term \cite{skoda}.
We understand navigation as a process of planning a path of a mobile robot from its current position to a desired goal location, following the planned path, and avoiding any discovered obstacles along the way \cite{fuzzyMPC}, \cite{NONMPC}. The desired paths have to
fulfill several conditions to ensure safety and feasibility of the navigation \cite{safety}. Moreover, the paths can be also compared in terms of desirability for example short or smooth paths are usually more desirable than long and curved ones. Such paths should therefore be preferred in the navigation process \cite{archi}. Beyond the path planning, the navigation problem also involves reacting to changes of the environment model. Robots are required to move towards target in a short time and avoid either static or dynamic obstacles observed by their sensors, which involves efficient path planning and valid obstacle avoidance \cite{MMPC}, \cite{autoMPC}, \cite{borelli}. Though these two topics have been well researched, currently, there is no ideal solution to handling the navigation problem within cluttered dynamic environments. The proposed algorithms were studied for the ATLASCAR project in which the group for Robotics and Automation at the University of Aveiro, has setup and adapted a common commercial vehicle to provide a versatile  framework  to  develop  studies  and  research
\cite{atlas}, \cite{atlascar1}, \cite{atlascar2}.
\section{Context of the Problem and \\  Proposed Solution}
The ATLACAR2 has sensors, such as lidar, that measure the distance to obstacles in front and around the vehicle. The obstacles can be static, such as a large pothole, or moving, such as a slow-moving vehicle. The most common maneuver from the driver is to temporarily move to another lane, drive past the obstacle, and move back to the original lane afterward. In this case, we want to design an obstacle avoidance system that moves the ATLASCAR2 around a moving obstacle in the lane using throttle and steering angle. This system uses an adaptive Model Predictive Controller that updates both the predictive model and the mixed input/output constraints at each control interval. Moreover, this vehicle is also equipped with a lane-keeping assist system: it has a sensor, such as camera or laser, that measures the lateral deviation and relative yaw angle between the centerline of a lane and the ATLASCAR2; it also measures the current lane curvature and its derivative. Depending on the curve length that the sensor can view, the curvature in front of the vehicle can be calculated from the current curvature and its derivative. This system keeps the autonomous car travelling along the centerline of the lanes on the road by adjusting the front steering angle. The goal for lane keeping control is to drive both lateral deviation and relative yaw angle close to zero.
\subsection{Theoretical background - Adaptive MPC}
Model Predictive Control is an advanced method that predicts future behavior using a linear-time-invariant (LTI) dynamic model. These predictions are never exact and a good strategy is to make MPC insensitive to prediction errors. If the plant is strongly nonlinear or its characteristics vary dramatically with time, MPC performance might become unacceptable because LTI prediction accuracy degrade. A method that can address this degradation by adapting the prediction model for changing operating conditions is called Adaptive MPC: this control strategy uses a fixed model structure, but allows the model parameters to evolve with time. Ideally, whenever the controller requires a prediction, it uses a model appropriate for the current conditions. At each control interval, the adaptive MPC controller updates the plant model and nominal conditions. Once updated, the model and conditions remain constant over the prediction horizon. The plant model used as the basis for the adaptive MPC must be an LTI discrete-time, state-space model with a structure as follows:

@@

where the matrices $A$, $B_u$, $B_v$, $B_d$, $C$, $D_v$ and $D_d$ can vary with time. The other parameters in the expression are:

@@

	\item $y$ are the plant outputs, including both measured and unmeasured. Is it necessary at least one measured output.

@@

In the adaptive MPC control there are additional requirements for the plant model as the sample time $T_s$ that has to be constant and identical to the MPC control interval. This control strategy prohibits direct feed-through from any manipulated variable to any plant output. Thus, $D_u = 0$ in the above model. Finally the input and output signal configuration remains constant.
A traditional MPC controller includes a nominal operating point at which the plant model applies, such as the condition at which you linearize a nonlinear model to obtain the LTI approximation. In adaptive MPC, as time evolves you should update the nominal operating point to be consistent with the updated plant model. It is possible to rewrite the plant model in terms of deviations from the nominal conditions:

@@

where the matrices $A$, $B$, $C$ and $D$ are updated with respect to time. The other parameters in the expression are:

@@

The adaptive MPC uses a Kalman filter to update its controller states and adjusts the two required gain matrices, $L$ and $M$, at each control interval to maintain consistency with the updated plant model. We are considering a linear-time-varying Kalman filter (LTVKF):
\label{eqn:LTVKF}
\begin{array}{lll}
L_k=\big(A_kP_{k|k-1}C_{m,k}^{T}+N\big)\big(C_{m,k}P_{k|k-1}C_{m,k}^{T}+R\big)^{-1}\\
M_k =P_{k|k-1}C_{m,k}^{T}\big(C_{m,k}P_{k|k-1}C_{m,k}^{T}+R\big)^{-1}\\
P_{k+1|k}=A_kP_{k|k-1}A_k^{T}-\big(A_kP_{k|k-1}C_{m,k}^{T}+N\big)L_k^{T}+Q.
where 
 \item $Q$, $R$ and $N$ are constant covariance matrices defined as in MPC state estimation;
 \item $A_k$ and $C_{m,k}$ are state-space parameter matrices for the entire controller state;
 \item $P_{k|k-1}$ is the state estimate error covariance matrix at time $k$ based on the information available at time $k-1$;
 \item $L_k$ and $M_k$ are the updated Kalmen filter gain matrices.
The gain and the state error covariance matrix depend upon the model parameters
and the assumptions leading to the constant $Q$, $R$ and $N$ matrices.
\section{Obstacle Avoidance}

@@

The collision avoidance problem is very dependent on the vehicle modeling since it is a requirement for  adaptive MPC law design. 
\begin{figure}[!h]

@@

@@

The model used in this paper should take into account the kinematic and dynamic aspects of the vehicle. Here, we present a non linear mathematical model of a vehicle used for the development of a collision avoidance system.
The model has four states and two inputs:
\textbf{x}=\begin{bmatrix}
   x&y&\theta&v 
  \end{bmatrix}^{T}
,\qquad 
\textbf{u}=\begin{bmatrix}
T&\delta 
\end{bmatrix}^{T}
where $(x,y)$ are the global coordinates of the center point of the car, $\theta$ is the heading angle of the car body with respect to the $x$-axis and $v$ is the speed of the car (positive). The manipulated variables are $T$ the throttle (positive when accelerating/negative when braking) and $\delta$ the steering angle ($0$ when aligned with car, counterclockwise positive).
\begin{figure}[!h]
	\includegraphics[width=0.70\columnwidth]{./figure/car_model.pdf}
	\caption{Bicycle model of a car.}
	\label{fig:car_model}
The ATLASCAR2 can be modeled using the non-linear kinematic bicycle model described by the following equations of motion:

@@

where $C_L$ is the car length. In order to simplify the model, it is assumed that only the front wheel can be steered. Moreover, in this paper it is assumed that the ATLASCAR2 does not slip, any slippage is thus considered as an external disturbance. Under this assumption, the slip angle is zero, meaning that the velocity is directed along the heading of the vehicle. In order to use MPC, the state space model needs to be linearized and also re-written in a more compact form:

@@

The simple linearized approximation of the system to describe the dynamics of the ATLASCAR2 will be evaluated at the operating conditions; the extended form of the considered model is the following:
	\label{eqn:dynamics_model_obstacle_avoidance_linearized}
	\dot{x} = -v\sin(\theta)\cdot\theta+\cos(\theta)\cdot v\\
	\dot{y} = v\cos(\theta)\cdot\theta+\sin(\theta)\cdot v\\
	\dot{\theta} =(\tan(\delta)/C_L)\cdot v+ \bigg(v\bigg(\tan(\delta)^2+1\bigg)/C_L\bigg)\cdot \delta\\
	\dot{v} =0.5\cdot T
The system to be controlled is usually modeled by a discrete state-space model in the MPC literature. Therefore, (\ref{eqn:dynamics_model_non_linear})

@@

where $A_d$ and $B d$ are the state and control matrices for the discrete state-space equation, respectively, which can be calculated with the Euler method as

@@

We designed a Model Predictive Controller that can make the ATLASCAR2 mantain a desired velocity of and stay in the middle of center lane. We used an Adaptive MPC controller because it handles the nonlinear vehicle dynamics more effectively than a traditional MPC controller; in fact, the latter uses a constant plant model but the former allows us to provide a new plant model at each control interval. Because the new model describes the plant dynamics more accurately at the new operating condition, an adaptive MPC controller performs better than a traditional MPC controller. In practice at each control interval, the adaptive MPC controller updates the plant model and the nominal conditions. Once updated, the model and the conditions remain constant over the prediction horizon. In path tracking using adaptive MPC, it is common to formulate the constrained control problem as a real-time optimization problem subject to hard constraints on plant variables and soft constraints on outputs; at the beginning, we specified the constraints for the manipulated variables: to prevent the ATLASCAR2 from accelerating or decelerating too quickly, we added an hard constraint on the throttle rate of change and another one on the steering angle rate of change. We used an approach that takes advantage of the ability of MPC to handle constraint explicitly. When an obstacle is detected, it defines an area on the road (in terms of constraints) that the ATLASCAR2 must not enter during the prediction horizon. At the next control interval the area is redefined based on the new positions of the vehicle and the obstacle until passing is completed.

@@

	\item the first constraint is an upper bound on the $y$-coordinate (right boundary of the road);
	\item the second constraint is a lower bound on the $y$-coordinate (left boundary of the road);
	\item the third constraint is for the obstacle avoidance; even though no obstacle is detected at the nominal condition, we must add this fake constraint here because we cannot change the dimensions of the constraint matrices at run time (it has the same form as the second one);

@@

where $W$ is the width of the road, $cI$ and $cS$ are the required parameters such that the ATLASCAR2 must be above the line formed from the vehicle to safe zone corner for left/right passing and finally $x_{\max}$ represents the position of the closest obstacle while $x_{\min}$ depicts the location of our vehicle. 

@@

simulation examples. We tried to choose parameters that were as close as possible to a real situation: the sampling time used in the discretization of the system is $T_s= 0.02$s while the values of the prediction and the control horizon are respectively $p_H=25$ and $c_H=5$. In all simulations, the distance between the front and rear axles is $C_L=5$m and the witdh of the vehicle is $C_W=2$m. The saturation ranges of the control inputs are: the steering angle lies in $\big[-\frac{\pi}{30}, +\frac{\pi}{30}\big]$rad/sec while in order to prevent the ATLASCAR2 from accelerating or decelerating too quickly, we impose an hard constraint of 2.5m$^2$/sec on the throttle rate of change. Moreover we are using a constant reference signal for the velocity of $v=20$m/s ($\simeq72$km/h).

@@

In this first simulation we assumed that the road is straight and has 3 lanes each one 4 meters wide. Moreover the ATLASCAR2 drives in the middle of the center lane while the road is completely free and when there is an obstacle the vehicle passes it only using the right lane (the same simulation can be launched so that the car goes over to the left fast lane). When the vehicle detects the obstacle, the constraints are computed: if the ATLASCAR2 is already in the adjacent lane, it uses the safety zone as the constraint, otherwise the vehicle must be above the line formed from the ATLASCAR2 to safe zone corner for right passing. If the vehicle is parallel to the obstacle, it  uses the safety zone as the constraint and finally if it has passed the obstacle, it uses the inactive constraint to go back to the center lane. In this first example the obstacle moves in the same direction as the vehicle.
\begin{figure}[!h]

@@

	\label{fig:obstacleAvoidance_start}

@@

	\label{fig:obstacleAvoidance_middle}

@@

	\label{fig:obstacleAvoidance_middle_end}

@@

	\label{fig:obstacleAvoidance_end}

@@

@@

@@

Consequently, additional obstacles were added to make the scenario more complex. The vehicle is capable of overtaking the obstacles to the right or left depending on their positions with respect to the road. If the $y$-coordinate of the closest obstacle is greater than 0, then the vehicle overtakes to the right, while if the $y$-coordinate is less than 0 the overtaking takes place on the left lane. We also hypothesized that the obstacles move at a different speeds but that they are initially at a common distance. In case two obstacles are too close during the simulation and their distance is less than the detection range which is $30$m, the ATLASCAR2 perceives the objects as a single entity and adapts to the situation. In this case the obstacles drive in the wrong direction with respect to the ATLASCAR2 (but the same results can be verified in the same direction).
\begin{figure}[!h]

@@

	\label{fig:obstacleAvoidance_random_2}

@@

	\label{fig:obstacleAvoidance_random_1}

@@

\subsubsection{Vehicle Braking and Obstacles Overtaking}
Finally we have improved the code related to the mixed Input/Output constraints so that in the case there are 3 obstacles that block the road and drive at a lower speed than the ATLASCAR2, the velocity of the vehicle decreases in order to prevent the collision. Here it is depicted a simulation in which there are the obstacles at the same $x$-coordinate. Two obstacles have a constant speed of $8$m/s while the one on the left lane has a velocity of $15$m/s. At the beginning the vehicle detects all the other cars on the road. In the first part of the simulation, the vehicle brakes because there is no enough space for overtaking the cars as shown in the Fig.\ref{fig:velocity_braking}. A collision will
happen if the vehicle continues to follow the initially planned trajectory with the reference velocity. It is possible to notice that the speed decreases because the applied throttle is negative, so a consistent deceleration is set after $\simeq1.5$s as depicted in Fig.\ref{fig:throttle_braking}. After few seconds the fastest obstacle moves and makes available the left lane. Dramatic changes of steering angle in early stage are observed in Fig.\ref{fig:delta_braking} and consequently also on the heading angle in Fig.\ref{fig:theta_braking}. Then, the ATLASCAR2 returns to the reference velocity during the overtaking of the two obstacles. The path of the vehicle is is depicted in Fig. \ref{fig:lateral_braking}. It is seen that the ATLASCAR2 avoids the  obstacles and returns to the road  centerline with a low overshoot.
%%% Simulation Braking
\begin{figure}[!h]
		\includegraphics[width=1\columnwidth]{../../MATLAB/three_obstacles_no_overtaking/figure/braking_0.pdf}
		\label{fig:braking_0}
		\includegraphics[width=1\columnwidth]{../../MATLAB/three_obstacles_no_overtaking/figure/braking_1.pdf}
		\label{fig:braking_1}
		\includegraphics[width=1\columnwidth]{../../MATLAB/three_obstacles_no_overtaking/figure/braking_2.pdf}
		\label{fig:braking_2}
		\includegraphics[width=1\columnwidth]{../../MATLAB/three_obstacles_no_overtaking/figure/braking_3.pdf}
		\label{fig:braking_3}
		\includegraphics[width=1\columnwidth]{../../MATLAB/three_obstacles_no_overtaking/figure/braking_4.pdf}
		\label{fig:braking_4}
		\includegraphics[width=1\columnwidth]{../../MATLAB/three_obstacles_no_overtaking/figure/braking_5.pdf}
		\label{fig:braking_5}
		\includegraphics[width=1\columnwidth]{../../MATLAB/three_obstacles_no_overtaking/figure/braking_6.pdf}
		\label{fig:braking_6}
	\caption{Simulation of braking and overtaking obstacles.}
\begin{figure}[!h]
	\includegraphics[width=1\columnwidth]{../../MATLAB/three_obstacles_no_overtaking/figure/VelocityVsTime.pdf}
	\caption{Velocity $v$ with respect to time.}
	\label{fig:velocity_braking}
\begin{figure}[!h]
	\includegraphics[width=1\columnwidth]{../../MATLAB/three_obstacles_no_overtaking/figure/ThrottleVsTime.pdf}
	\caption{Throttle $T$ with respect to time.}
	\label{fig:throttle_braking}
\begin{figure}[!h]
	\includegraphics[width=1\columnwidth]{../../MATLAB/three_obstacles_no_overtaking/figure/SteeringAngleVsTime.pdf}
	\caption{Steering angle $\delta$ with respect to time.}
	\label{fig:delta_braking}
\begin{figure}[!h]
	\includegraphics[width=1\columnwidth]{../../MATLAB/three_obstacles_no_overtaking/figure/HeadingAngleVsTime.pdf}
	\caption{Heading angle $\theta$ with respect to time.}
	\label{fig:theta_braking}
\begin{figure}[!h]
	\includegraphics[width=1\columnwidth]{../../MATLAB/three_obstacles_no_overtaking/figure/LateralPositionVsTime.pdf}
	\caption{Lateral position $y$ with respect to time.}

>>>>

\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage{listings}
\usepackage{stfloats}
\usepackage{midfloat}
\usepackage{color}
\usepackage[table]{xcolor}
\usepackage{lipsum}
\usepackage[utf8x]{inputenc}
\usepackage{float}
\usepackage{bbm}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[colorlinks=true,linkcolor=black,anchorcolor=black,citecolor=black,filecolor=black,menucolor=black,runcolor=black,urlcolor=black]{hyperref}

\newtheorem{thm}{Theorem}[section]
\newtheorem{defn}[thm]{Def.}
\newtheorem{propen}[thm]{Prop.}
\newtheorem{cor}[thm]{Cor.}
\renewcommand\IEEEkeywordsname{Keywords}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
		T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\title{Short-term Path Planning with Collision Avoidance in Autonomous Driving based on Adaptive MPC}
\author{
	\IEEEauthorblockN{Alberto Franco}
	\IEEEauthorblockA{Department of Information Engineering,\\ Università degli Studi di Padova, Italy\\		 \href{mailto:alberto.franco.3@studenti.unipd.it}{alberto.franco.3@studenti.unipd.it}\\
	\href{mailto:alberto.franco@ua.pt}{alberto.franco@ua.pt}}
	\and
	\IEEEauthorblockN{Vitor Santos}
	\IEEEauthorblockA{Department of Mechanical Engineering,\\ IEETA, Universidade de Aveiro, Portugal\\ \href{mailto:vitor@ua.pt}{vitor@ua.pt}}}

\begin{document}
\maketitle
\begin{abstract}
	In this paper we present a different strategy for a self-driving car short-term path planning in space-based approach. The aim of this project is to study and implement a path tracking framework in order to prevent ATLASCAR2 collision with moving obstacle vehicles. The proposed algorithm, based on the Model Predictive Control paradigm, solves an optimal problem formulated in terms of cost minimization under constraints. Results demonstrate and verifies the feasibility and the usefulness of the method considering different scenarios.
\end{abstract}

\begin{IEEEkeywords}
	autonomous driving, Model Predictive Control, obstacle avoidance.
\end{IEEEkeywords}

\section{Introduction}
In robotic research, the problem of navigation is among the most important. Basically, all autonomous mobile robots need some kind of navigation to fulfill the mobile term \cite{skoda}.
We understand navigation as a process of planning a path of a mobile robot from its current position to a desired goal location, following the planned path, and avoiding any discovered obstacles along the way \cite{fuzzyMPC} \cite{NONMPC}. The desired paths have to fulfill several conditions to ensure safety and feasibility of the navigation \cite{safety}. Moreover, the paths can be also compared in terms of specifications for example short or smooth paths are usually more desirable than long and curved ones. Such paths should therefore be preferred in the navigation process \cite{archi}. Beyond the path planning, the navigation problem also involves reacting to changes of the environment model. Robots are required to move towards the target in a short time and avoid either static or dynamic obstacles observed by their sensors, which involves efficient path planning and valid obstacle avoidance \cite{MMPC} \cite{autoMPC} \cite{borelli}. Though these two topics have been well researched, currently, there is no ideal solution to handling the navigation problem within cluttered dynamic environments.
\section{Context of the Problem and \\  Proposed Solution}
The proposed algorithm was studied for the ATLASCAR project in which the group for Robotics and Automation at the University of Aveiro, has setup and adapted a common commercial vehicle to provide a versatile  framework  to  develop  studies  and  research \cite{atlas} \cite{atlascar1} \cite{atlascar2}.
In Fig.\ref{fig:atlascar2} is reported the chosen platform which is an electric car, a Mitsubishii-MiEV of 2015 with a range of a $100$km.
The fact that the vehicle is electric allows to use the energy stored in the batteries, making it easier to power the sensors installed.
\begin{figure}[H]
	\centering
	\includegraphics[width=0.70\columnwidth]{./figure/atlascar2.jpg}
	\vspace{-12pt}
	\caption{ATLASCAR2, based on an electric car, a Mitsubishii-MiEV}
	\label{fig:atlascar2}
\end{figure}

The ATLACAR2 has sensors, such as lidar, that measure the distance to obstacles in front and around the vehicle. The obstacles can be static, such as a large pothole, or moving, such as a slow-moving vehicle. The most common maneuver from the driver is to temporarily move to another lane, drive past the obstacle, and move back to the original lane afterward. In this case, we want to design an obstacle avoidance system that moves the ATLASCAR2 around a moving obstacle in the lane using throttle and steering angle. This system uses an adaptive Model Predictive Controller that updates both the predictive model and the mixed input/output constraints at each control interval.
\section{Theoretical background\\ Adaptive Model Predictive Control}
Model Predictive Control is an advanced method that predicts future behavior using a linear-time-invariant (LTI) dynamic model. These predictions are never exact and a good strategy is to make MPC insensitive to prediction errors. If the plant is strongly nonlinear or its characteristics vary dramatically with time, MPC performance might become unacceptable because LTI prediction accuracy degrade \cite{mpctoolbox}. A method that can address this degradation by adapting the prediction model for changing operating conditions is called Adaptive MPC: this control strategy uses a fixed model structure, but allows the model parameters to evolve with time. Ideally, whenever the controller requires a prediction, it uses a model appropriate for the current conditions. At each control interval, the adaptive MPC controller updates the plant model and nominal conditions. Once updated, the model and conditions remain constant over the prediction horizon. The plant model used as the basis for the adaptive MPC must be an LTI discrete-time, state-space model with a structure as follows:

@@

where the matrices $A$, $B_u$, $B_v$, $B_d$, $C$, $D_v$ and $D_d$ can vary with time. The other parameters in (\ref{eqn:Adaptive_MPC_plant_discrete}) are:

@@

	\item $y$ are the plant outputs, including both measured (necessary at least one) and unmeasured.

@@

In the adaptive MPC control, there are additional requirements for the plant model, like the sample time $T_s$ that has to be constant and identical to the MPC control interval. This control strategy prohibits direct feed-through from any manipulated variable to any plant output. Thus, $D_u = 0$ in the above model.
A traditional MPC controller includes a nominal operating point at which the plant model applies, such as the condition at which you linearize a nonlinear model to obtain the LTI approximation (equilibrium, reference trajectory and the	
most updated value) \cite{mpctoolbox}. In adaptive MPC, as time evolves you should update the nominal operating point to be consistent with the updated plant model. It is possible to rewrite the plant model in terms of deviations from the nominal conditions:

@@

where the matrices $A$, $B$, $C$ and $D$ are updated with respect to time. The other parameters in (\ref{eqn:Adaptive_MPC_nominal_condition}) are:

@@

The adaptive MPC uses a Kalman filter to update its controller states which include the plant, the disturbance and measurement noise model states.
\section{Moving Obstacle Avoidance}

@@

The collision avoidance problem is very dependent on the vehicle modeling since it is a requirement for  adaptive MPC law design. In Fig.\ref{fig:obstacleAvoidance} is depicted a typical scenario of overtaking of a moving obstacle.
\begin{figure}[H]

@@

	\vspace{-25pt}

@@

The model used in this paper should take into account the kinematic and dynamic aspects of the vehicle. Here, we present a non linear mathematical model of a vehicle used for the development of a collision avoidance system.
The model has four states and two inputs:
\[ 
\textbf{x}=\begin{bmatrix}
   x&y&\theta&v 
  \end{bmatrix}^{T}
,\qquad 
\textbf{u}=\begin{bmatrix}
T&\delta 
\end{bmatrix}^{T}
\]
where $(x,y)$ are the global coordinates of the center point of the car, $\theta$ is the heading angle of the car body with respect to the $x$-axis and $v$ is the linear speed of the car (positive). The manipulated variables are $T$ the throttle (positive when accelerating/negative when braking) and $\delta$ the steering angle ($0$ when aligned with car, counterclockwise positive). Fig.\ref{fig:car_model} shows a schematic sketch of the applied nonlinear bicycle model and the related parameters.
\begin{figure}[!h]
	\centering
	\includegraphics[width=0.70\columnwidth]{./figure/car_model.pdf}
	\vspace{-12pt}
	\caption{Bicycle model of a car.}
	\label{fig:car_model}
\end{figure}

The ATLASCAR2 can be modeled using the non-linear kinematic bicycle model described by the following equations of motion \cite{safety} \cite{swarms}:

@@

where $C_L$ is the car length. In order to simplify the model, it is assumed that only the front wheel can be steered. Moreover, in this paper it is assumed that the ATLASCAR2 does not slip, so any slippage is thus considered as an external disturbance. Under this assumption, the slip angle is zero, meaning that the velocity is directed along the heading of the vehicle. In order to use MPC, the state space model needs to be linearized with a first order approximation and also re-written in a more compact form:

@@

The simple linearized approximation of the system to describe the dynamics of the ATLASCAR2 will be evaluated at the operating conditions. The system to be controlled is usually modeled by a discrete state-space model in the MPC literature. Therefore, (\ref{eqn:dynamics_model_non_linear})

@@

where $A_d$ and $B_d$ are the state and control matrices for the discrete state-space equation, respectively, which can be calculated with the Euler method as

@@

We designed a Model Predictive Controller that can make the ATLASCAR2 mantain a desired velocity and stay in the middle of center lane. We used an Adaptive MPC controller because it handles the nonlinear vehicle dynamics more effectively than a traditional MPC controller; in fact, the latter uses a constant plant model but the former allows us to provide a new plant model at each control interval. Because the new model describes the plant dynamics more accurately at the new operating condition, an adaptive MPC controller performs better than a traditional MPC controller. In practice, at each control interval, the adaptive MPC controller updates the plant model and the nominal conditions. Once updated, the model and the conditions remain constant over the prediction horizon. In motion planning using adaptive MPC, it is common to formulate the constrained control problem as a real-time optimization problem subject to hard constraints on plant variables and soft constraints on outputs; at the beginning, we specified the constraints for the manipulated variables: to prevent the ATLASCAR2 from accelerating or decelerating too quickly, we added an hard constraint on the throttle rate of change and another one on the steering angle rate of change. We used an approach that takes advantage of the ability of MPC to handle constraint explicitly. In Fig.\ref{fig:flowchart} is depicted a conditional state machine designed for higher-level behavior planning:
\begin{figure}[H]
	\includegraphics[width=1\columnwidth]{./figure/flowchart/flowchart.pdf}
	\vspace{-20pt}
	\caption{Behaviour planning conditional flowchart.}
	\label{fig:flowchart}

When an obstacle is detected, it defines an area on the road (in terms of constraints) that the ATLASCAR2 must not enter during the prediction horizon. At the next control interval the area is redefined based on the new positions of the vehicle and the obstacle until passing is completed.

@@

	\item the first constraint is an upper bound on the $y$-coordinate (left boundary of the road);
	\item the second constraint is a lower bound on the $y$-coordinate (right boundary of the road);
	\item the third constraint is for the obstacle avoidance; even though no obstacle is detected at the nominal condition, we must add this fake constraint here because we cannot change the dimensions of the constraint matrices at run time;

@@

where $W$ is the width of the road, $cI$ and $cS$ are the required parameters such that the ATLASCAR2 must be above the line formed from the vehicle to safe zone corner for left/right passing and finally $x_{\max}$ represents the position of the closest obstacle while $x_{\min}$ depicts the location of our vehicle. Fig.\ref{fig:constraint} shows a schematic sketch of the constraints that are computed at each sampling time $T_s$ in the case of a left overtaking.
\begin{figure}[H]
	\includegraphics[width=1\columnwidth]{./figure/constraint/constraint.pdf}
	\vspace{-20pt}
	\caption{Constraints in the case of left overtaking.}
	\label{fig:constraint}

@@

simulation examples. We tried to choose parameters that were as close as possible to a real situation: the sampling time used in the discretization of the system is $T_s= 0.02$s while the values of the prediction and the control horizon are respectively $p_H=25$ and $c_H=5$. In all simulations, the distance between the front and rear axles is $C_L=5$m and the witdh of the vehicle is $C_W=2$m. The saturation ranges of the control inputs are: the steering angle lies in $\big[-\frac{\pi}{30}, +\frac{\pi}{30}\big]$rad/sec while in order to prevent the ATLASCAR2 from accelerating or decelerating too quickly, we impose an hard constraint of 2.5m/sec$^2$ on the throttle rate of change. Moreover we are using a constant reference signal for the velocity of $v=20$m/s ($\simeq72$km/h).

@@

In this first simulation
(Fig.\ref{fig:obstacleAvoidance_one_obstacle}) we assumed that the road is straight and has 3 lanes of 4 meters wide each. Moreover the ATLASCAR2 drives in the middle of the center lane while the road is completely free and when there is an obstacle, the vehicle passes it only using the right lane (the same simulation can be launched so that the car goes over to the left fast lane). When the vehicle detects the obstacle, the constraints are computed: if the ATLASCAR2 is already in the adjacent lane, it uses the safety zone as the constraint, otherwise the vehicle must be above the line formed from the ATLASCAR2 to safe zone corner for right passing. If the vehicle is parallel to the obstacle, it  uses the safety zone as the constraint and finally if it has passed the obstacle, it uses the inactive constraint to go back to the center lane. In this example the obstacle moves in the same direction as the vehicle.
\begin{figure}[H]

@@

@@

@@

@@

@@

	\vspace{-20pt}

@@

	\label{fig:obstacleAvoidance_one_obstacle}

@@

Consequently, additional obstacles were added to make the scenario more complex (Fig.\ref{fig:obstacleAvoidance_random}). The vehicle is capable of overtaking the obstacles to the right or left depending on their positions with respect to the road. If the $y$-coordinate of the closest obstacle is greater than 0, then the vehicle overtakes to the right, while if the $y$-coordinate is less than 0, the overtaking takes place on the left lane. We also hypothesized that the obstacles move at a different speeds but that they are initially at a common distance. In case two obstacles are too close during the simulation and their distance is less than the detection range which is $30$m, the ATLASCAR2 perceives the objects as a single entity and adapts to the situation. In this case the obstacles drive in the opposite direction with respect to the ATLASCAR2 (but the same results can be verified in the same direction).
\begin{figure}[H]

@@

@@

@@

\subsubsection{Vehicle Braking and Obstacles Overtaking}
Finally we have improved the code related to the mixed Input/Output constraints so that in the case there are 3 obstacles that block the road and drive at a lower speed than the ATLASCAR2, the velocity of the vehicle decreases in order to prevent the collision. In Fig.\ref{fig:braking} is depicted a simulation in which there are the obstacles at the same $x$-coordinate. Two obstacles have a constant speed of $8$m/s while the one on the left lane has a velocity of $15$m/s.
\begin{figure}[H]
	\centering
	\begin{minipage}[t]{\columnwidth}
		\includegraphics[width=1\columnwidth]{../../MATLAB/three_obstacles_no_overtaking/figure/braking_0.pdf}
	\end{minipage}
	\begin{minipage}[t]{\columnwidth}
		\includegraphics[width=1\columnwidth]{../../MATLAB/three_obstacles_no_overtaking/figure/braking_1.pdf}
	\end{minipage}
	\begin{minipage}[t]{\columnwidth}
		\includegraphics[width=1\columnwidth]{../../MATLAB/three_obstacles_no_overtaking/figure/braking_2.pdf}
	\end{minipage}
	\begin{minipage}[t]{\columnwidth}
		\includegraphics[width=1\columnwidth]{../../MATLAB/three_obstacles_no_overtaking/figure/braking_3.pdf}
	\end{minipage}
	\begin{minipage}[t]{\columnwidth}
		\includegraphics[width=1\columnwidth]{../../MATLAB/three_obstacles_no_overtaking/figure/braking_4.pdf}
	\end{minipage}
	\begin{minipage}[t]{\columnwidth}
		\includegraphics[width=1\columnwidth]{../../MATLAB/three_obstacles_no_overtaking/figure/braking_5.pdf}
	\end{minipage}
	\begin{minipage}[t]{\columnwidth}
		\includegraphics[width=1\columnwidth]{../../MATLAB/three_obstacles_no_overtaking/figure/braking_6.pdf}
	\end{minipage}	
	\vspace{-20pt}
	\caption{Simulation of braking and overtaking obstacles.}
	\label{fig:braking}
\end{figure}
